% 1.  Cook's Distance 1977
% 2.  Extension to 
% 3.  Demidenko 
% 4.  Schabenburger
%---------------------------------------------------------------% 

\begin{abstract}
The purpose of this article is to review the use of diagnostic measures for LME models.
\end{abstract}


%---------------------------------------------------------------%
\section{Schabenberger: SUMMARY AND CONCLUSIONS}

Standard residual and inﬂuence diagnostics for linear models can be extended to linear mixed models. The dependence of ﬁxed-effects solutions on the covariance parameter estimates has important ramiﬁcations in perturbation analysis. To gauge the full impact of a set of observations on the analysis, covariance parameters need to be updated, which requires reﬁtting of the model. The experimental INFLUENCE option of the MODEL statement in the MIXED procedure (SAS 9.1) enables you to perform iterative and noniterative inﬂuence analysis for individual observations and sets of observations.

The conditional (subject-speciﬁc) and marginal (population-averaged) formulations in the linear mixed model enable you to consider conditional residuals that use the estimated BLUPs of the random effects, and marginal residuals which are deviations from the overall mean. Residuals using the BLUPs are useful to diagnose whether the random effects components in the model are speciﬁed correctly, marginal residuals are useful to diagnose the ﬁxed-effects components. Both types of residuals are available in SAS 9.1 as an experimental option of the MODEL statement in the MIXED procedure.

It is important to note that inﬂuence analyses are performed under the assumption that the chosen model is correct. Changing the model structure can alter the conclusions. Many other variance models have been ﬁt to the data presented in the repeated measures example. You need to see the conclusions about which model component is affected in light of the model being ﬁt. For example, modeling these data with a random intercept and random slope for each child or an unstructured covariance matrix will affect your conclusions about which children are inﬂuential on the analysis and how this inﬂuence manifests itself.
%-----------------------------------------------------------------%
\section{Diagnostic Methods for OLS models}
% Cook's Distance for OLS models
% http://www.amstat.org/meetings/jsm/2012/onlineprogram/AbstractDetails.cfm?abstractid=305411
Influence diagnostics are formal techniques allowing for the identification of observations that exert substantial 
influence on the estimates of fixed effects and variance covariance parameters. 

The idea of influence diagnostics for a given observation is to quantify the effect of omission of this observation 
from the data on the results of the model fit. To this aim, the concept of likelihood displacement is used. 

%---------------------------------------------------------------%
We have developed a function in R, which allows performing influence diagnostics for linear mixed effects models 
fitted using the lme() function from the nlme package. 

The use of the new function is illustrated using data from a randomized clinical trial.

%---------------------------------------------------------------%

\subsection{Influence Diagnostics: Basic Idea and Statistics} %1.1.2
%http://support.sas.com/documentation/cdl/en/statug/63033/HTML/default/viewer.htm#statug_mixed_sect024.htm

The general idea of quantifying the influence of one or more observations relies on computing parameter estimates based on all data points, removing the cases in question from the data, refitting the model, and computing statistics based on the change between full-data and reduced-data estimation. 


%---------------------------------------------------------------%

\subsection{Influence Analysis for LME Models} %1.1.3
The linear mixed effects model is a useful methodology for fitting a wide range of models. However, linear mixed effects models are known to be sensitive to outliers. \citet{CPJ} advises that identification of outliers is necessary before conclusions may be drawn from the fitted model.

Standard statistical packages concentrate on calculating and testing parameter estimates without considering the diagnostics of the model.The assessment of the effects of perturbations in data, on the outcome of the analysis, is known as statistical influence analysis. Influence analysis examines the robustness of the model. Influence analysis methodologies have been used extensively in classical linear models, and provided the basis for methodologies for use with LME models.
Computationally inexpensive diagnostics tools have been developed to examine the issue of influence \citep{Zewotir}.
Studentized residuals, error contrast matrices and the inverse of the response variance covariance matrix are regular components of these tools.

\subsection{Influence Statistics for LME models} %1.1.4
Influence statistics can be coarsely grouped by the aspect of estimation that is their primary target:
\begin{itemize}
\item overall measures compare changes in objective functions: (restricted) likelihood distance (Cook and Weisberg 1982, Ch. 5.2)
\item influence on parameter estimates: Cook's  (Cook 1977, 1979), MDFFITS (Belsley, Kuh, and Welsch 1980, p. 32)
\item influence on precision of estimates: CovRatio and CovTrace
\item influence on fitted and predicted values: PRESS residual, PRESS statistic (Allen 1974), DFFITS (Belsley, Kuh, and Welsch 1980, p. 15)
\item outlier properties: internally and externally studentized residuals, leverage
\end{itemize}
